---
title: "Palmerpenguins Dataset"
Owner: 'Patrick Dorante'
output: html_document
---

# Working with data with R programming language
# Step # 1
1) Import your dataset to rStudio. Above the Global Environment click 'import Dataset'
2) If datasets are packages so  Be sure to install your packages with this function #install.packages() 
3) If datasets are already installed to your computer so used these functions # library(), data() or attach() to load the dataset



Practice the palmerpenguins dataset for data exploration and visualization 
Load palmerpenguins packages to work on a new project and import tidyverse
The goal of this project is to find out number of species, 
where are they located? and sex etc... including data visualization.

# Manipulate data using the tidyverse package

```{r setup, include=FALSE}
library(tidyverse)
library(palmerpenguins)
data('penguins')
library(here)
library(skimr)
library(janitor)
library(dplyr)
knitr::opts_chunk$set(echo = TRUE)
```

# Overview of the data using summary() and glimpse() methods

```{r palmerpenguins}
summary(penguins)
glimpse(penguins)
head(penguins)
str(penguins)
colnames(penguins)
```

# Preview of the data using head() methods
## Using functions such as str() or colnames() to check for data structures

1) head(penguins)
2) str(penguins)
3) colnames(penguins)


# Creating tibble as a result to return a neatly organized table that could print and view easily.
 
as_tibble(penguins)

# The readr and readxl packages

The readr package let users read rectangular data. readr supports several read_functions:

1. read_csv(): comma-separated values (.csv) files

2. read_tsv(): tab-separated values files

3. read_delim(): general delimited files

4. read_fwf(): fixed-width files

5. read_table(): tabular files where columns are separated by white-space

6. read_log(): web log file

7. read_excel()

# Example how to read a .csv file and .xlsx
read_csv(readr_example("palmerpenguins.csv"))

read_excel(readxl_example("palmerpenguins.xlsx"))

8 core tidyverse packages for data analysis and and visualization
# ggplot
# tibble
# tidyr
# readr
# purrr
# dplyr 
# stringr
# forcats

Explore data with R packages may not be too complex

# Step #2 
# Data cleaning with R

Step #2 consists of how to clean your data

# Let's Cleaning up data with dplyr and janitor packages and commands as parts of the analysis
These functions will give a summary of what is in the data and cleaning
1. skim_without_charts()
2. rename()
3. rename_with()
4. select()
5. clean_names()
6. glimpse()
7. remove_empty()
8. remove_constant()


# %>% is a pipe operator called 'and then'. 
R takes %>% as a long continuous line

if you want to work with some specific variables or columns used the select() command
# rename() and rename_with() let you rename precise variable

```{r penguins}
select(penguins, species, island, sex, year)

# Without pipe operator use select() function to work with one group of variable
select(penguins, species) # 1st. method

penguins %>% 
  select(species) # 2nd method


# Work with one or two or more groups of variables
select(penguins, species, island, sex, year)

penguins %>% 
  select(species, island, sex, year) # 2nd method


# Work without a group of variable or column
select(penguins, -species)

penguins %>% 
  select(-species) # 2nd method

# Work without two or more groups of variable or column
select(penguins, -species, -sex)

penguins %>% 
  select(species, island, sex, year) # 2nd method


# Create a variable and assign it to a new select function using %>% pipe operator
land_penguins <- penguins %>% 
                 select(species, island, sex, year)

# Output the new variable land_penguins
land_penguins

# rename() function would change a column's name
penguins %>% rename(BMI = body_mass_g)

# Assign this function to a variable
bmi <- penguins %>% 
  rename(BMI = body_mass_g, Annual = year)

# Output the variable
bmi

# Change column names to upper case and be more consistent
rename_with(penguins, toupper)

# Change column names to lower case
rename_with(penguins, tolower)

# clean_names() 
##Returns names with only lowercase letters, with _ as a separator
#Handles special characters and spaces
#Appends numbers to duplicated names
#Converts “%” to “percent” to retain meaning

penguins %>% 
  clean_names() %>% head()

```

# Step # 3
# Organize Data
 can change a variable and create a new variable:
1. arrange()
2. summarize()
3. filter()
4. group_by()
5. drop_na()
6. max() min()
7. mean()


```{r}
# Filtering data 
penguins %>% 
  select(species, island, sex, year) %>%
  filter(species == 'Chinstrap')

# The arrange() function lets you sort a dataframe by a variable of interest
penguins %>% 
  arrange(flipper_length_mm) %>%
  head()

# DESCENDANT ORDER
penguins %>% 
  arrange(desc(flipper_length_mm)) %>%
  rename(flipperLength = flipper_length_mm) %>%
  head()


# drop_na() function drop NA or remove missing values
penguins %>% 
  arrange(desc(flipper_length_mm)) %>%
  drop_na() %>%
  group_by(species) %>%
  head()

# mean() function group of columns
penguins %>% 
  arrange(desc(species)) %>%
  drop_na() %>%
  group_by(species) %>%
  summarise(Average_body_mass = mean(bill_length_mm)) 

# max() function group of columns
penguins %>% 
  arrange(desc(flipper_length_mm)) %>%
  drop_na() %>%
  group_by(species) %>%
  summarise(Max_body_mass = max(body_mass_g))

# min() function group of columns
penguins %>% 
  arrange(desc(flipper_length_mm)) %>%
  drop_na() %>%
  group_by(species) %>%
  head(Min_body_mass = min(body_mass_g))

```
# Step # 4
# Tranform Data
can change a variable and create a new variable:
1. mutate()
2. unite()
3. separate()





You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.



